API design for machine learning software: experiences from the scikit-learn project

### 2013-09-10 paper.tex
more copyediting/trimming

The \sklearn project \citep{pedregosa2011} provides an open source machine learning
library for the Python programming language. The ambition of the project is to
provide efficient and well-established machine learning tools within a
programming environment that is accessible to non-machine learning experts and
reusable in various scientific areas. The project is not a novel domain-specific
language, but a library that provides machine learning idioms to a
general-purpose high-level language. Among other things, it includes classical learning
algorithms, model evaluation and selection tools, as well as
preprocessing procedures. The library is
distributed under the simplified BSD license, encouraging its use in both
academic and commercial settings.

\sklearn is a library, i.e.\ a collection of classes
and functions that users import into Python programs. Using \sklearn therefore
requires basic Python programming knowledge. No command-line interface, let
alone a graphical user interface, is offered for non-programmer users. Instead,
interactive use is made possible by the Python interactive interpreter, and its
enhanced replacement IPython \citep{perez2007ipython}, which offer a
\textsc{matlab}-like working environment specifically designed for scientific
use.

@@

Started in 2007, \sklearn is developed by an international team of over a dozen
core developers, mostly researchers from various fields (e.g.,
computer science, neuroscience, astrophysics). The project also benefits
from many occasional contributors proposing small bugfixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Documentation and examples are required for all features,
and major changes must pass code review by at least two
developers not involved in the implementation of the proposed change.

@@

\citet{pedregosa2011} briefly presented \sklearn and
benchmarked it against several competitors.
In this paper, we instead present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api} the central application
programming interface (API) and then describe, in section~\ref{sec:advanced-api},
advanced API mechanisms built on the core interface.
Section~\ref{sec:implementation} briefly describes the implementation.
Section~\ref{sec:comparison} then
compares \sklearn to other major projects in terms of API\@.
Section~\ref{sec:future_work} outlines some of the objectives for
a \sklearn 1.0 release.
We conclude by summarizing the major points of this paper in
section~\ref{sec:conclusions}.

@@

All objects within \sklearn share a uniform common basic API consisting of three
complementary interfaces: an \textit{estimator} interface for building and
fitting models, a \textit{predictor} interface for making predictions and a
\textit{transformer} interface for converting data. In this section, we describe
these three interfaces, after reviewing our general principles and data
representation choices.

@@

As much as possible, our
design choices have been guided so as to avoid the proliferation of framework
code. We try to adopt simple conventions and to limit to a minimum the number of
methods an object must implement. The API is designed to adhere to the following
broad principles:

@@

In \sklearn, the data representation is kept as close as possible
to the matrix representation: datasets are encoded as
two-dimensional NumPy arrays or SciPy sparse matrices.
While these may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as the ones used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
on efficient NumPy and SciPy vectorized operations while keeping
the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices.
From a practical point of view, these formats also provide a collection of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

@@

The \textit{estimator} interface is at the core of the
library. It defines instantiation mechanisms of objects and exposes a
\texttt{fit} method for learning a model from training data.  All supervised and
unsupervised learning algorithms (e.g., for classification, regression or
clustering) are offered as objects implementing this interface. Machine
learning tasks like feature extraction, feature selection or dimensionality
reduction are also provided as estimators.

Estimator initialization and actual learning are strictly separated,
in a way that is similar to partial function application:
an estimator is initialized from a set of named constant hyper-parameter values
(e.g., the $C$ constant in SVMs)
and can be considered as a function
that maps these values to actual learning algorithms.
The constructor of an estimator does not see any actual data, nor does it perform any actual learning.
All it does is attach the given parameters to the object.
For the sake of convenient model inspection, hyper-parameters are set as public attributes,
which is especially important in model selection settings.
For ease of use, default hyper-parameter values are also provided
for all built-in estimators.
These default values are set to be relevant in many common
situations in order to make estimators as effective as possible
\textit{out-of-box} for non-experts.

@@

clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"} for
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C},
the strength of the regularization) are not explicitly given and
thus set to the default values. Upon calling \texttt{fit}, a model is
learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
and stored within the object for later use.
%% (In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name);
to build a random forest on
the same data, one would simply replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by
\texttt{RandomForestClassifier()}.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection, feature extraction and dimensionality reduction
algorithms are all provided as transformers within the library.  In our example,
to standardize the input \texttt{X\_train} to zero mean and unit variance
before fitting the logistic regression estimator,
one would write:
\begin{pythoncode}
from sklearn.preprocessing import StandardScaler

@@

Some machine learning algorithms are expressed naturally
as meta-algorithms parametrized on simpler algorithms.
Examples include ensemble methods which
build and combine several simpler models (boosting, bagging, random forests),
or multiclass classification schemes which combine multiple binary classifiers.
In \sklearn, such algorithms are implemented as \textit{meta-estimators}.
They take as input an existing estimator
and use it internally for learning and making predictions.
All meta-estimators implement the regular estimator interface.
% XXX bagging example here

Since meta-estimators require users to construct nested objects,
the decision to implement a meta-estimator
rather than integrate the behavior it implements
into existing estimators classes
is always based on a trade-off between generality and ease of use.

@@

For each hyper-parameter combination and each train/validation split
generated by the cross-validation scheme, \texttt{GridSearchCV}
and \texttt{RandomizedSearchCV} fit their base estimator on the training set and
evaluate its performance on the validation set.  In the end, the best performing
model on average is retained and exposed as the public attribute
\texttt{best\_estimator\_}.

@@

clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
clf.fit(X_train, y_train)
y_pred = clf.predict(X_test)
\end{pythoncode}
\caption{Optimizing the hyper-parameters of an SVM for $F_1$ score.}
\label{svm-gridsearch}
\end{figure*}

The code in Figure~\ref{svm-gridsearch} illustrates how to find
hyper-parameter settings for an SVM classifier (SVC)
that maximize $F_1$ score
through 10-fold cross-validation on the training set.
Two distinct hyper-parameter grids are
considered for the linear and radial basis function (RBF) kernels;
an SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
so using a single parameter grid would waste computing time
trying out effectively equivalent settings.
Additionally, we see that
\texttt{GridSearchCV} has a \texttt{predict} method, just like any other classifier:
it delegates the \texttt{predict}, \texttt{predict\_proba}, \texttt{transform} and
\texttt{score} methods to the best estimator
(optionally after re-fitting it on the whole training set).

@@

To ease code reuse, simplify implementation and skip the introduction of
superfluous classes, the Python principle of \textit{duck typing} is exploited
throughout the codebase. This means that estimators are defined by interface,
not by inheritance, where the interface is entirely implicit
as far as the programming language is concerned.
Duck typing allows both for extensibility and
flexibility: as long as an estimator follows the API and conventions
outlined in Section~\ref{sec:core-api}, then it can be used in lieu of a
built-in estimator (e.g., it can be plugged into pipelines or grid search)
and external developers are not forced to inherit from any \sklearn class.

In other places of the library, in particular in the vectorization code
for unstructured input, the toolkit is also designed to be
extensible. Here, estimators provide hooks for user-defined code: objects or
functions that follow a specific API can be given as arguments at vectorizer
construction time. The library then calls into this code,
communicating with it by passing objects of standard Python/NumPy types,
and keeping it agnostic of the \sklearn class hierarchy.

Our rule of thumb is that user code should not be tied to \sklearn---which is a
\textit{library}, and not a framework. This principle avoids a
common problem with object-oriented design, which is that users wanting a
``banana'' should not get ``a gorilla holding the banana and the entire jungle''
(J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).
Programs using \sklearn should not be intimately tied to it,
so that their code can be reused with other toolkits or in other contexts.

@@

To facilitate the installation of \sklearn,
the set of external dependencies is kept to a bare minimum:
only Python, NumPy and SciPy are required for a functioning installation.
Binary distributions of these are available for the major platforms.
Visualization functionality uses on Matplotlib \citep{hunter2007matplotlib}
and Graphviz \citep{gansner2000},
but neither is required to perform machine learning or prediction.
When feasible, external libraries are integrated into the codebase.
In particular, \sklearn includes modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!
and wrapped using Cython.

@@

Other existing machine learning packages
such as SofiaML\footnote{\url{https://code.google.com/p/sofia-ml}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/~vw}}
are intended to be used as command-line tools
(and sometimes do not offer any kind of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
users will find that they still need programming to process input/output,
and will do so in a variety of languages.
By contrast, \sklearn allows users to implement that entire workflow
in a single program, written in a single language,
and developed in a single working environment.
This also makes it easier for researchers and developers
to exchange and collaborate on software, as dependencies and setup are kept to a
minimum.

@@

\section{Future directions}
\label{sec:future_work}
There are several directions that the \sklearn project
aims to focus on in future development.
At present, the library does not support some classical machine learning
algorithms,
including neural networks, ensemble meta-estimators for
bagging or subsampling strategies and missing value completion algorithms.
However, tasks like structured prediction or reinforcement learning are
considered out of scope for the project,
since they would require quite different data representations and APIs.

At a lower-level, parallel processing is a potential point of improvement.
Some estimators in \sklearn are already able to leverage multicore processors,
but only in a coarse-grained fashion.
At present, parallel processing is difficult to accomplish in the Python environment;
\sklearn targets the main implementation, CPython,
which cannot execute Python code on multiple CPUs simultaneously.
It follows that any parallel task decomposition must either be done
inside Cython modules,
or at a level high enough to warrant the overhead
of creating multiple OS-level processes,
and the ensuing inter-process communication.
Parallel grid search is an example of the latter approach
which has already been implemented.
Recent versions of Cython include support for the OpenMP standard
\citep{dagum1998openmp},
which is a viable candidate technology
for more fine-grained multicore support in \sklearn.

@@

We have discussed the \sklearn API
and the way it maps machine learning concepts and tasks
onto objects and operations in the Python programming language.
We have shown how a consistent API across the package makes \sklearn
very \textbf{usable} in practice: experimenting with different learning
algorithm is as simple as substituting a new class definition.
Through composition interfaces such as Pipelines, Feature Unions,
and meta-estimators, these simple building blocks lead to an API which is
\textbf{powerful}, and can accomplish a wide variety of learning tasks
within a small amount of easy-to-read code.
Through duck-typing, the consistent API leads to a library that is
easily \textbf{extensible}, and allows user-defined estimators to be
incorporated into the \sklearn workflow without any explicit object
inheritance.

While part of the \sklearn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.
The power, and extensibility of the \sklearn API is evidenced
by the large and growing user-base, its use to solve real
problems across a wide array of fields,
as well as the appearance of third-party packages
that follow the \sklearn conventions. Examples of such packages include
\textit{astroML}\footnote{\url{http://astroml.org}}
\citep{vanderplas2012astroML}, a package providing
machine learning tools for astronomers, and
\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
implementation. The source code of
the recently-proposed sparse multiclass
algorithm of \citet{mblondel-mlj2013}, released as
part of the
\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
package, also follows the \sklearn conventions.
To maximize ease of use, we encourage more researchers
to follow these conventions when releasing their software.

>>>>

The \sklearn project \citep{pedregosa2011} provides an open source machine learning
library for the Python programming language. The ambition of the project is to
provide efficient and well-established machine learning tools within a
programming environment that is accessible to non-machine learning experts and
reusable in various scientific areas. The project is not a novel domain-specific
language, but a library that provides machine learning idioms to a
general-purpose high-level language. It includes classical learning algorithms,
model evaluation and selection tools, as well as preprocessing procedures.
The library is distributed under the simplified BSD license,
encouraging its use in both academia and industry.

\sklearn is a library, i.e.\ a collection of classes
and functions that users import into Python programs.
No command-line interface, let alone a graphical user interface,
is offered for non-programmer users. Instead,
interactive use is made possible by the Python interactive interpreter, and its
enhanced replacement IPython \citep{perez2007ipython}, which offers a
\textsc{matlab}-like working environment especially suited for scientific use.

@@

Since 2007, \sklearn has been developed by an international team of more than a dozen
core developers, mostly researchers from various fields (e.g.,
computer science, neuroscience, astrophysics). The project also benefits
from many occasional contributors proposing small bugfixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Documentation and examples are required for all features,
and major changes must pass code review by at least two
developers not involved in the implementation of the proposed change.

@@

\citet{pedregosa2011} briefly presented \sklearn and
benchmarked it against several competitors.
In this paper, we present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api} the central application
programming interface (API) and then describe, in section~\ref{sec:advanced-api},
advanced API mechanisms built on the core interface.
Section~\ref{sec:implementation} briefly describes the implementation.
Section~\ref{sec:comparison} then
compares \sklearn to other major projects in terms of API\@.
Section~\ref{sec:future_work} outlines some of the objectives for
a \sklearn 1.0 release.
We conclude by summarizing the major points of this paper in
section~\ref{sec:conclusions}.

@@

All objects in \sklearn share a uniform basic API consisting of three
complementary interfaces: an \textit{estimator} interface for building and
fitting models, a \textit{predictor} interface for making predictions and a
\textit{transformer} interface for converting data. In this section, we describe
these three interfaces, after reviewing our general principles and data
representation choices.

@@

As much as possible, \sklearn has been designed in such a way
as to avoid the proliferation of framework code.
We adopt simple conventions and limit the number of methods
an object must implement to a minimum.
The API is designed to adhere to the following principles:

@@

In \sklearn, the data representation is kept as close as possible
to the matrix representation: datasets are encoded as
two-dimensional NumPy arrays or SciPy sparse matrices.
While these may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as the ones used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
on efficient NumPy and SciPy vectorized operations while keeping
the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices.
These formats also provide a collection of
data loading and conversion tools which make them very easy to use in many
contexts. For tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

@@

The \textit{estimator} interface is at the core of the
library. It defines instantiation mechanisms of objects and exposes a
\texttt{fit} method for learning a model from training data.  All supervised and
unsupervised learning algorithms (e.g., for classification, regression or
clustering) are offered as objects implementing this interface. Machine
learning tasks like feature extraction and selection are also provided as estimators.

Estimator initialization and actual learning are strictly separated,
in a way that is similar to partial function application:
an estimator is initialized from a set of named hyper-parameter values
(e.g., the $C$ constant in SVMs)
and can be considered a function
that maps these values to actual learning algorithms.
The constructor does not see any actual data.
All it does is attach the given parameters to the object.
For the sake of model inspection, hyper-parameters are set as public attributes,
which is especially important in model selection settings.
Default hyper-parameter values are provided for all built-in estimators.
These defaults are set to be relevant in many common
situations in order to make estimators effective \textit{out-of-box}.

@@

clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"} for
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C},
the strength of the regularization) are not explicitly given and
thus set to the default values. Upon calling \texttt{fit}, a model is
learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
and stored on the object for later use.
%% (In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name);
to build a random forest on
the same data, one would simply replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by
\texttt{RandomForestClassifier()}.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection, feature extraction and dimensionality reduction
algorithms are all provided as transformers.  In our example,
to standardize the input \texttt{X\_train} to zero mean and unit variance
before fitting the logistic regression estimator,
one would write:
\begin{pythoncode}
from sklearn.preprocessing import StandardScaler

@@

Some machine learning algorithms are expressed naturally
as meta-algorithms parametrized on simpler algorithms.
Examples include ensemble methods which
build and combine several simpler models (boosting, bagging, random forests),
or multiclass classification schemes which combine multiple binary classifiers.
In \sklearn, such algorithms are implemented as \textit{meta-estimators}.
They take as input an estimator which is used
internally for learning and making predictions.
All meta-estimators implement the regular estimator interface.
% XXX bagging example here

Since meta-estimators require users to construct nested objects,
the decision to implement a meta-estimator
rather than integrate the behavior it implements
into existing estimator classes
is always based on a trade-off between generality and ease of use.

@@

For each hyper-parameter combination and each train/validation split
generated by the cross-validation scheme, \texttt{GridSearchCV}
and \texttt{RandomizedSearchCV} fit their base estimator on the training set and
evaluate its performance on the validation set.  In the end, the best performing
model on average is retained and exposed as the public attribute
\texttt{best\_estimator\_} (optionally re-fit on the whole training set),
but \texttt{GridSearchCV} also has a \texttt{predict} method,
which delegates to the best model's \texttt{predict},
so the CV object is itself an estimator and a predictor.

@@

clf = GridSearchCV(SVC(), param_grid, scoring="f1", cv=10)
clf.fit(X_train, y_train)
\end{pythoncode}
\caption{Optimizing the hyper-parameters of an SVM for $F_1$ score.}
\label{svm-gridsearch}
\end{figure*}

The code in Figure~\ref{svm-gridsearch} shows how to find
hyper-parameter settings for an SVM classifier (SVC)
that maximize $F_1$ score
through 10-fold cross-validation on the training set.
Two distinct hyper-parameter grids are
considered for the linear and radial basis function (RBF) kernels;
an SVM with a linear kernel accepts a $\gamma$ parameter, but ignores it,
so using a single parameter grid would waste computing time
trying out effectively equivalent settings.

@@

The Python principle of \textit{duck typing} is exploited
throughout the codebase. This means that estimators are defined by interface,
not by inheritance, where the interface is entirely implicit
as far as the programming language is concerned.
Duck typing allows both for extensibility and
flexibility: as long as an estimator follows the API and conventions
outlined in Section~\ref{sec:core-api}, then it can be used in lieu of a
built-in estimator (e.g., in pipelines and grid search)
and developers are not forced to inherit from any \sklearn class.

In other places of the library, in particular in the vectorization code
for unstructured input, the toolkit is also designed to be
extensible. Here, estimators provide hooks for user-defined code: objects or
functions that follow a specific API can be given as arguments at vectorizer
construction time. The library then calls into this code,
communicating with it using objects of standard Python/NumPy types,
and keeping it agnostic of the \sklearn class hierarchy.

Our rule of thumb is that user code should not be tied to \sklearn---which is a
\textit{library}, not a framework. We try to avoid a
common problem with object-oriented design: that users wanting a
``banana'' instead get ``a gorilla holding the banana and the entire jungle''
(J.~Armstrong, cited by \citealp[p.~213]{seibel2009coders}).
Programs using \sklearn should not be intimately tied to it,
so that their code can be reused or in other contexts.

@@

To facilitate the installation of \sklearn,
the set of external dependencies is kept to a bare minimum:
only Python, NumPy and SciPy are required for a functioning installation.
Binary distributions of these are available for the major platforms.
Visualization functionality uses Matplotlib \citep{hunter2007matplotlib}
and Graphviz \citep{gansner2000},
but neither is required to perform machine learning or prediction.
When feasible, external libraries are integrated into the codebase.
In particular, \sklearn includes modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!
and wrapped using Cython.

@@

Other machine learning packages,
such as SofiaML\footnote{\url{https://code.google.com/p/sofia-ml}}
and Vowpal~Wabbit,\footnote{\url{http://hunch.net/~vw}}
are often designed as command-line tools
and may not offer any kind of API\@.
While this has the advantage that users are not tied
to a particular programming language,
they will find that they still need programming to process input/output,
and will do so in a variety of languages.
By contrast, \sklearn allows users to implement that entire workflow
in a single program, written in a single language.
This also makes it easier for researchers and developers
to exchange and collaborate on software, as dependencies and setup are kept to a
minimum.

@@

\section{Future directions}
\label{sec:future_work}
There are several directions that the \sklearn project
aims to focus on in future development.
At present, the library does not support some classical machine learning
algorithms,
including neural networks and subsampling strategies.
However, tasks like structured prediction or reinforcement learning are
considered out of scope for the project,
since they would require quite different data representations and APIs.

At a lower-level, parallel processing is a potential point of improvement.
Some estimators in \sklearn are already able to leverage multicore processors,
but only in a coarse-grained fashion.
Parallel processing is difficult to accomplish in the Python environment;
\sklearn targets the main implementation, CPython,
which cannot execute Python code on multiple CPUs simultaneously.
It follows that any parallel task decomposition must either be done
inside Cython modules,
or at a level high enough to warrant the overhead
of creating multiple OS-level processes,
and the ensuing inter-process communication.
Parallel grid search is an example of the latter approach
which has already been implemented.
Recent versions of Cython include support for the OpenMP standard
\citep{dagum1998openmp},
which is a viable candidate technology
for more fine-grained multicore support in \sklearn.

@@

We have discussed the \sklearn API
and the way it maps machine learning concepts and tasks
onto objects and operations in the Python programming language.
We have shown how a consistent API across the package makes \sklearn
easy to use: experimenting with different learning
algorithms can be as simple as substituting a class name.
Through composition interfaces such as pipelines and meta-estimators,
simple building blocks lead to an API which is powerful,
and can accomplish a wide variety of learning tasks
within a small amount of code.
Through duck-typing, the consistent API leads to an extensible library
where custom algorithm can be incorporated into the workflow without inheritance.

While part of the \sklearn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.
The power and extensibility of the \sklearn API is evidenced
by the large and growing user-base, its use to solve real
problems across a wide array of fields,
as well as the appearance of third-party packages
that follow \sklearn conventions. Examples of such packages include
\textit{astroML}\footnote{\url{http://astroml.org}}
\citep{vanderplas2012astroML}, a machine learning toolkit for astronomy,
and \textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
implementation. The source code of the recently-proposed sparse multiclass
algorithm of \citet{mblondel-mlj2013}, released as part of the
\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
package, also follows the \sklearn conventions.
We encourage researchers to follow these conventions when releasing software.
